---
title: "MetaAnalysis Assignment"
output: html_notebook
author: "Renée Claes and Maxime Keutgen"
---
## Introduction
In this notebook, we try to reproduce the Meta-Analysis of Hornsey & al 2016.
We start by loading libraries.
```{r}
library(haven)
library(metafor)
library(tidyverse)
NEP <- read_sav("NEP.sav")
ObjKnow      <- read_sav("Objective Knowledge.sav")
SubjKnow <- read_sav("Subjective Knowledge.sav")
FreeMark <- read_sav("Support for Free-market Ideology.sav")
TrustScientist <- read_sav("Trust in Scientists.sav")
```

## Objective knowledge
First, we will start with one dataset, objective knowledge. 
A bit of datacleaning is required (shortening name of one study). 

```{r}
ObjKnow[15,1] <- "Hine, D. W., Reser et al (2013)" 

```
Also, we need an estimate of the variance of the effect size. Since these are correctional studies, the correlations coefficients have been transformed to Fisher Z scores$^1$ for inferential purposes. 
The correctional coefficient is given by :
$$ r = \frac{cov(X,Y)}{\sigma_X\sigma_Y} $$
Fisher transformation is given by :
$$ z = \frac{1}{2} ln \bigg(\frac{1+r}{1-r} \bigg) $$
If $X_1..X_n$ and $Y_1,..,Y_n$ are i.i.d paired data, then $z$ is asymptotically normally distributed$^2$ :
$$ z \sim \mathcal{N} \Bigg(\frac{1}{2}ln\bigg(\frac{1+r}{1-r}\bigg),\frac{1}{N-3} \Bigg) $$
We can then compute the variance of the Fisher Z-scores.
```{r}
ObjKnow <- ObjKnow %>% mutate(var=1/(ObjKnow$Sample_size-3 ))
```
We now have at our disposal point estimate of the effect size (Z scores) and of the variance of this effect size. We start by fitting a Random Effect model (because it is standard and I don't know what else to do at this stage).
```{r}
REM <- rma(yi = ObjKnow$FishersZ,vi = ObjKnow$var,method = "REML")
```
Let's look at fancy plots.
```{r}
plot.rma.uni(REM)

```




## References
[1] Fisher, R. A. (1915). "Frequency distribution of the values of the correlation coefficient in samples of an indefinitely large population". Biometrika. 10 (4): 507–521. doi:10.2307/2331838. hdl:2440/15166. JSTOR 2331838
[2] Fisher, R. A. (1921). "On the 'probable error' of a coefficient of correlation deduced from a small sample" (PDF). Metron. 1: 3–32.
